---
title: "Homework starter document"
format:
  html: 
    toc: true # includes a table of contents in rendered html format
execute: # for all code chunks
    warning: false # do not display warnings
    message: false # do not display messages
---

# Set up

## reading in packages

```{r}
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

# quick look at data 
str(drought_exp)
class(drought_exp)
```

## cleaning

```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```


#Set Up Models

## 0. Null model

```{r}
model0 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame
```

## 1. total biomass as a function of SLA, water treatment, and species

```{r}
# saturated model
model1 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)

par(mfrow = c(2, 2))
plot(model1)

```


## 2. total biomass as a function of SLA and water treatment

```{r}
model2 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)

plot(model2)
```


## 3. total biomass as a function of SLA and species

```{r}
model3 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)

plot(model3)
```


## 4. total biomass as a function of water treatment and species

```{r}
model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)

plot(model4)

```

# Model selection

From the `MuMIn` package

```{r}
model.sel(model0,
          model1, 
          model2, 
          model3,
          model4)
```

# Model summary

```{r}
summary(model4)
```

# Model predictions

```{r}
model_preds <- ggpredict(model1, 
                         terms = c("water_treatment", 
                                   "species_name"))
```

# Visualization

```{r}
# plot function within ggeffects
plot(model_preds, # model predictions
     limit_range = TRUE, # limit the range of predictions to the range of predictor values
     show_data = TRUE) + # show the underlying data
  # everything below this is ggplot() stuff
  theme_classic() + # classic theme
  labs(title = "Preliminary model visualization") + # plot title
  theme(panel.grid = element_blank())  # getting rid of gridlines
```

The "more customizable" way:

```{r}
# creating new data frame of model predictions for plotting
model_preds_for_plotting <- model_preds %>% 
  rename(sla = x, # renaming columns to make this easier to use
         water_treatment = group,
         species_name = facet)

# use View(model_preds_for_plotting) 
# to compare this to the original model_preds data frame

ggplot() +
  # underlying data
  geom_point(data = drought_exp_clean,
             aes(x = sla,
                 y = total_g,
                 color = water_treatment)) +
  # model prediction 95% CI ribbon
  geom_ribbon(data = model_preds_for_plotting,
              aes(x = sla, 
                  y = predicted,
                  ymin = conf.low,
                  ymax = conf.high,
                  fill = water_treatment),
              alpha = 0.2) +
  # model prediction lines
  geom_line(data = model_preds_for_plotting,
            aes(x = sla, 
                y = predicted,
                color = water_treatment)) +
  # cleaner theme
  theme_classic() +
  # creating different panels for species
  facet_wrap(~species_name) 
```


# Making a table

```{r}
# one option for a single model
flextable::as_flextable(model4) 

# another option for a single model
gtsummary::tbl_regression(model4)

# comparing models
modelsummary::modelsummary( # this function takes a list of models
  list( 
    "null" = model0, # "model name" = model object
    "model 1" = model1,
    "model 2" = model2,
    "model 3" = model3,
    "model 4" = model4
  )
) 


```


a. Make a table or list of all the models from class and the last one you constructed on your own. Write a caption for your table. (8 points)

```{r}
model_table <- read_csv("model_table.csv")

ft <- flextable::as_flextable(model_table) |> 
  set_caption(caption = "Table 1. Water treatment and species names the best predictors for this model. This table shows the four models we have; water treatment + species name, saturated (all three), sla + species name, sla + water treatment, and the null (no predictors).")

ft
```

Caption: table captions typically go above the table. Number the table and provide a title. Describe what is in the table (columns and rows).

Table: In your table, each row should be a model with the model number (1, 2, 3, etc.) and the predictors for each model.


b. Write a 5-6 sentence “statistical methods” section. (8 points)
Your answer should be in paragraph form and include:

I addressed the central question to examine the influence of specific leaf area, water treatment, and species on total plant mass. I did this using 5 models to see which combination of predictors describes the variation in plant mass. To determine the model that best described total plant mass I input all the models into a model selection table which showed the intercept, predictors in that model, the AICc and delta. Using the AIC to compare different versions of a model with variations in the predictors variables I chose the model with the lowest AIC which shows that the linear model predicts the data best. I visually determined that the final model conformed to the assumptions of a linear model, to evaluate the linear model assumptions ran the code to make a linear model with the predictions for model 4. I think visually confirmed that the Residuals vs Fitted and Scale- Location plots were normal, even, and linear; that the QQ Residuals plot was normal and followed for the most part a straight diagonal line; and that Constant Leverage: Residuals vs. Factor Levels didn't have an outliers that would alter the data modeling.

c. Make a visualization of the model predictions with underlying data for your “best” model. (20 points)
Show and annotate all your code. For full credit:

```{r}
ggpredict(model4, 
          terms = c("water_treatment", 
                    "species_name")) |> 
  plot(show_data = TRUE) + # show the underlying data
  #scale_color_manual(values = c(Encelia californica, Echscholzia californica, Grindelia camporum, Acmispon glaber, Nasella pulchra, Penstemon centranthifolius, Salvia leucophylla)) +
  #scale_fill_manual(values = c(Well, Drought)) +
  theme_classic() + # classic theme
  labs(title = "Model 4 Predictions Visualization") + # plot title
  theme(panel.grid = element_blank())  # getting rid of gridlines


```

# Visualization

**Note:** this is the quick way to visualize model predictions. For your homework, you may find it easier to plot things yourself (i.e. not using the `plot()` functionality within `ggeffects`). See lecture for an example of how to do this.  

```{r}
# plot function within ggeffects

```

make the underlying data more transparent than the model predictions
display species names in full (not as species codes like ENCCAL or ESCCAL)
display water treatment types in full (not as WW or DS)
represent well-watered and drought stressed treatments with different colors
use colors that are not the default ggplot() colors
facet your plot by species
remove the legend
finalize the plot

d. Write a caption for your visualization. (6 points)
Include a data citation.

e. Write a 3-4 sentence results section. (10 points)
Your answer should be in paragraph form and address the following points:

what predictors “best” described total mass (include model statistics here)?
on average, what differences did you find between water treatments?
on average, what differences did you find between species?


a. Revisit and summarize (6 points)
**What are the statistical tests the authors are using to address their main research question?**

The statistical tests that the authors use to address their main research question are Student t test or bivariate Pearson correlation analyses and linear regression models. 

**Insert the figure or table you described in Homework 2 here.**

b. Visual clarity (10 points)
**How clearly did the authors visually represent their statistics in figures? For example, are the x- and y-axes in a logical position? Do they show summary statistics (means and SE, for example) and/or model predictions, and if so, do they show the underlying data?**

I don't think the authors were very clear in their visual representation of the statistical analysis. The figure is not on an x-y coordinate it is a model showing the predictors showing regression beta weights and correlation coefficients.The figure does not show SE or CIs. It doesn't show underlying data just thicker arrows for predictors that have a higher coefficient.

c. Aesthetic clarity (10 points)
**How well did the authors handle “visual clutter”? How would you describe the the data:ink ratio?**

The figure is not very cluttered, it is actually pretty simple. I would say there should be more data shown, it in my opinion over simplifies the statistical analysis. The data to ink ratio is like 1:5 the only significant part of the figure is the coefficients showing how significant the predictor variables are.

d. Recommendations (can be longer than 4 sentences, 10 points)
**What recommendations would you make to make the figure better? What would you take out, add, or change? Provide explanations/justifications for each of your recommendations.**

I think this figure could be much more informational with some small changes. I would recommend showing the underlying data with SE and CI to demonstrate how significant, and confident the research is in there findings of which predictors are correlated to stigmatization. I think if the figure was on an x y coordinate and there was more than one figure to show the data from all the predictors it would be a better figure.


